---
title: "Assignment 1"
author: '10179889'
date: "09/02/2022"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    font:family: Lato
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

For this assignment, we have been provided with two different datasets. Using the knowledge we have gained from the workshops combined with further reading and wider resources, I will:

- Wrangle and tidy the data
- Summarise the data
- Visualise the data
- Build and interpret the appropriate models

## Packages

First, let's load in our packages using the `library()` function:

```{r}
library(tidyverse)
library(lme4)
library(lmerTest)
library(performance)
library(visdat)
library(ggthemes)
library(showtext)
```

The `{tidyverse}` package contains a collection of open source R packages that gives us access to a variety of useful functions and commands for data wrangling and visualisation. The `{showtext}` package enables us to load in and use over 1291 fonts from [Google Fonts](https://fonts.google.com/) for our data visualisations, whilst `{ggthemes}` contains extra themes for our plots. 


The `{afex}`and`{emmeans}` packages will allow us to build ANOVA models and perform post-hoc analyses, respectively. We'll be using the `{afex]` package instead of the `aov()` function in base R as it allows us to build ANOVA models that will work for our experimental designs, as well as using type III sums of squares by default. Finally, `{visdat}` will help us to spot any missing data in our datasets.

After downloading the **Lato** font from [Google Fonts](https://fonts.google.com/), 